|

### 2023-02-07 review/response.Rmd

---
title: "Response to review: New and simplified manual controls for projection and slice tours, with application to exploring classification boundaries in high dimensions"
author: "Ursula Laa, Alex Aumann, Dianne Cook, German Valencia"
date: "2023-01-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

 XXX If use new slice center control for fig 7, need to make a note of it here, updated the mathematica notebook and check text and captions correct.

 XXX References need to be consistent, all initials or all full names, and all sentence case or title case on titles. - Done

 XXX Check that slicing explanation mentions scaling the data! - where should this go?

@@

Summary

This paper describes a method for interacting with multidimensional data using slice tours in Mathematica (as well as a reduced-feature implementation in R).

Major Issues

- Reproducibility:
    - Code should be reproducible and not depend on user paths.
    - Data files should be uploaded along with code files (they don't seem to be part of supplemental material?)
    - Not having the ability to actually run the code (because the data wasn't included) makes it somewhat difficult to assess the software side of this paper.

\textcolor{blue}{Done. Previously we provided the public github repo URL so that all files to reproduce could be downloaded. Here we have included all the files in the supplementary zip. XXX TODO Please check the list of Supplementary material in the paper.}
\textcolor{red}{I think it is complete, but maybe can be structured better, to discuss}

@@

- Grammar: There are a lot of minor grammatical issues with this paper. I suggest utilizing a tool such as Grammarly to spot and fix them. The ones that I noticed as I read through the first section are detailed in minor issues below, but I am trying to focus more on the content than the style, so I will try to not nit-pick through the rest of the paper and trust that the authors can use the online tools.

\textcolor{blue}{Done. Apologies for the grammatical errors.}

- Cite your software packages appropriately - they're tools you depend on to do your work and should be cited in the same way papers are cited.

\textcolor{blue}{Done. They were all cited, but not, by our error, at the first mention.}

@@

\textcolor{blue}{This is better explained.}

  – Page 8, line 11: "The most relevant inbuilt Mathematica function for our purpose is the LocatorPane. This creates a region on the screen where the position of the mouse is captured and then converted to input that updates the graphics functions, enabling the manual navigation of the tour."

  - Finally, after all of this time, this is the one input that might be a bit harder to implement in R/Shiny, though I think via WebGL it should be workable. This is where the focus should be, rather than the previous two paragraphs, which are a distraction.

\textcolor{blue}{The first two paragraphs have been removed and the third restructured to explain why this is difficult in R/Shiny/rgl.}  
\textcolor{red}{Maybe have a single reply to all these points? It seems the blue text color in the paper is not there for all changes (paragraph on shiny is fully new?).}

### 2023-02-07 paper/appendix.Rmd
Getting ready to re-submit

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq.~\ref{matrices}. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee the orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuous and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalization.

>>>>

For $d=2$ projections, the projection matrix is the sub-matrix $A$, of $O$ (an orthonormal basis for the p-dimensional space), formed by its first two columns as illustrated in Eq. \ref{matrices}. Whereas orthonormality of the basis for the p-dimensional space is given by $e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}$, orthonormality of the projection matrix is expressed as $\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}$. Movement of the cursor takes the two components ${a_{m1},a_{m2}}$ into a selected new value ${a^*_{m1},a^*_{m2}}$. Although the motion is constrained by $a^{*2}_{m1}+a^{*2}_{m2}\leq 1$, this is not sufficient to guarantee the orthonormality of the new projection matrix. One possible algorithm to achieve this is
\begin{enumerate}
\item Cursor movement takes ${a_{m1},a_{m2}}\to {a^*_{m1},a^*_{m2}}$, called $V_m$ above.
\item The freedom to change the components $a_{m3}\cdots a_{mp}$ (the columns of $O$ not corresponding to the projection matrix $A$) is used to select a new orthonormal basis as follows:
\begin{enumerate}
\item For row $m$ one chooses $a^*_{m3}=\sqrt{1-a^{*2}_{m1}-a^{*2}_{m1}},~a^*_{m,k>3}=0$
\item For other rows, $a_{i\neq m,j\geq3}$ random selections in the range $(-1,1)$ are made.
\item The Gram-Schmidt algorithm is then used to obtain an orthonormal basis taking $e_m$ as the first vector (which is already normalized), and then $e_1\to e_1-(e_1\cdot e_m) e_m$, $e_1\to e_1/(e_1\cdot e_1)$, etc, resulting in $A^{**}$.
\end{enumerate}
\item this results in the orthonormal basis $O^*$ and a new projection matrix with $A^{**}_{m1}=a^*_{m1},~A^{**}_{m2}=a^*_{m2}$.
\end{enumerate}
The random completion of $O$ outside the projection matrix provides an exploration of dimensions orthogonal to the projection plane. However, it renders projections in a way that is not continuous and may be distracting. This can be alleviated by replacing the random completion with a rule restricting the size of the jumps, for example in step (b) the values $a_{i\neq m,j\geq3}$ can be left unchanged before orthogonalization.

### 2023-02-07 paper/appendix.tex
Getting ready to re-submit

For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
of \(O\) (an orthonormal basis for the p-dimensional space), formed by
its first two columns as illustrated in
Eq.\textasciitilde{}\ref{matrices}. Whereas orthonormality of the basis
for the p-dimensional space is given by
\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
projection matrix is expressed as
\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
to guarantee the orthonormality of the new projection matrix. One
possible algorithm to achieve this is

>>>>

For \(d=2\) projections, the projection matrix is the sub-matrix \(A\),
of \(O\) (an orthonormal basis for the p-dimensional space), formed by
its first two columns as illustrated in Eq. \ref{matrices}. Whereas
orthonormality of the basis for the p-dimensional space is given by
\(e_i\cdot e_j=\delta_{ij},{i,j,=1,\cdots, p}\), orthonormality of the
projection matrix is expressed as
\(\sum_{k=1}^p a_{ki} a_{kj}=\delta_{ij}, ~{i,j=1,2}, ~{k=1,\cdots,p}\).
Movement of the cursor takes the two components \({a_{m1},a_{m2}}\) into
a selected new value \({a^*_{m1},a^*_{m2}}\). Although the motion is
constrained by \(a^{*2}_{m1}+a^{*2}_{m2}\leq 1\), this is not sufficient
to guarantee the orthonormality of the new projection matrix. One
possible algorithm to achieve this is

### 2023-02-07 paper/paper.Rmd
Getting ready to re-submit

<!-- Background -->
From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, \textcolor{blue}{methods which work for arbitrary dimensions have more general utility}. A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*. Tours involve views of high-dimensional data with low-dimensional projections. In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*. (Various videos of tours, e.g. @dataviewer, @gt-pp-video, on the @ASA22, illustrate this concept.) Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections. (A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allow one to quickly see many, many different projections. After Asimov, there have been many tour developments, which are summarized in @lee2021.

@@

1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
2. Change values in row $m$, for example, if $d=2$ gives 
$$
A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
                              \vdots & \vdots \\
                              a^*_{m1} & a^*_{m2}\\
                              \vdots & \vdots \\
                              a_{p1} & a_{p2} 
        \end{array}\right].
$$ 
\noindent A large change in these values would correspond to making a large jump from the current projection. Small changes would correspond to tracking a cursor, making small jumps from the current projection.
3. Orthonormalise $A^*$, using Gram-Schmidt. 
    i. Normalise $\boldsymbol{a}^*_1$ and $\boldsymbol{a}^*_2$.
    ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1$.

@@

\textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing. However, the `plotly` package is designed to add to `ggplot2` graphics, not to be a programmable interactive graphics system. It lacks the capability to implement the primary tool needed for a manual tour, which is the tracking of the cursor position in small increments. In theory, the historical `locator()` function of base R should provide the necessary capability, but its behavior is not consistent between graphical devices. The cursor position control is possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary dimension of tours. The level of interactivity would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that currently operates across platforms.  What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement the resulting ideas as an animation in the `tourr` [@tourr] package. There is hope that better interactivity will be available for R soon, perhaps in javascript [e.g. @langevitour, @detourr], which might allow for the experimentation to develop new techniques.

@@

The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group). After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualization appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice center. The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided. It is worth noting that this zoom works best with data that has been centered and scaled. The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration into a later study, as shown in the applications below. More details about the implementation and usage instructions are given in the Appendix.

@@

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data. The three species are reasonably different in size, with Gentoo distinguished from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat + theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

@@

We start with a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in the top row of Figure \ref{slice1p}. \textcolor{blue}{The star glyph at right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables. This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.

@@

Mathematica provided a useful sandbox to experiment with the ideas presented in this paper. Most of the data, though, is first constructed using R, especially the classification boundaries. If new technology for interactive graphics with precise cursor control becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.

@@

- The Mathematica source code defining the new functions in mmtour.wl.
- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
- Appendix with additional details about manual controls, and Mathematica functions.
- The `Rmd` file that generated the paper, containing the R code to produce Figure 1.
- Data files for the examples.
- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.

>>>>

<!-- Background -->
From a statistical perspective 3D is a rare data dimension, so unlike in most 3D rotation computer graphics applications, \textcolor{blue}{methods which work for arbitrary dimensions have more general utility}. A good approach is to show projections from an arbitrary dimensional space to create dynamic data visualizations called *tours*. Tours involve views of high-dimensional data with low-dimensional projections. In his original paper on the grand tour, @As85 provided several algorithms for tour paths that could theoretically show the viewer the data *from all sides*. \textcolor{blue}{(Various videos of tours, e.g.} @dataviewer, @gt-pp-video, on the @ASA22, \textcolor{blue}{illustrate this concept.)} Prior to Asimov's work, there were numerous preparatory developments including @tukey's PRIM-9. PRIM-9 had user-controlled rotations on coordinate axes, allowing one to manually tour through low-dimensional projections. (A video [@PRIM9-video] illustrating the capabilities is available through \textcolor{blue}{the} video library of @ASA22.) Steering through all possible projections is impossible, unlike Asimov's tours which allow one to quickly see many, many different projections. After Asimov, there have been many tour developments, which are summarized in @lee2021.

@@

1. Provide $A$, and $m$. (Note that $m$ could also be automatically chosen as the component that is closest to the cursor position.)
2. Change values in row $m$, for example, if $d=2$ 
$$
A^* = [ \boldsymbol{a}^*_1~\boldsymbol{a}^*_2 ] = \left[ \begin{array}{cc} a_{11} & a_{12}\\
                              \vdots & \vdots \\
                              a^*_{m1} & a^*_{m2}\\
                              \vdots & \vdots \\
                              a_{p1} & a_{p2} 
        \end{array}\right].
$$ 
\noindent A large change in these values would correspond to making a large jump from the current projection. Small changes would correspond to tracking a cursor, making small jumps from the current projection.
3. Orthonormalise $A^*$, using Gram-Schmidt. 
    i. Normalise $\boldsymbol{a}^*_1$ and $\boldsymbol{a}^*_2$.
    ii. $\boldsymbol{a}^*_2 = \boldsymbol{a}^*_2 - (\boldsymbol{a}^*_1\cdot\boldsymbol{a}^*_2)\boldsymbol{a}^*_1$.

\textcolor{red}{a2 should be normalised after the second step? Fig 1: this is now showing tracking, with the yellow dot moving in each image?}

@@

\textcolor{blue}{Within R, the natural implementation for an interactive interface would use} `shiny` [@shinym] with `plotly` [@plotly] drawing. However, the `plotly` package is designed to add to `ggplot2` [@ggplot2] graphics, not to be a programmable interactive graphics system. It lacks the capability to implement the primary tool needed for a manual tour, which is the tracking of the cursor position in small increments. In theory, the historical `locator()` function of base R should provide the necessary capability, but its behavior is not consistent between graphical devices. The cursor position control is possible with `rgl` [@rgl] but this package is focused on 3D, not the arbitrary dimension of tours. The level of interactivity would have been possible with the `cranvas` system [@cranvas] or `rJava` [@rJava] but neither has a graphics toolkit that currently operates across platforms.  What we have chosen to do is to use the interactivity of Mathematica to experiment, and to implement the resulting ideas as an animation in the `tourr` [@tourr] package. There is hope that better interactivity will be available for R soon, perhaps in javascript [e.g. @langevitour, @detourr], which might allow for the experimentation to develop new techniques.

\textcolor{red}{plotly does 3D graphs that can be dragged as well, maybe need to clarify further? Are there any references that document shortcomings? Maybe this paragraph should be a subsection?}

@@

The examples presented in the application section below (and attached in notebook format in the supplementary material) all use our primary new function, `SliceDynamic`. This function typically accepts grouped data in the form of a matrix where the second last column details the name of the group and the last column details the group index (which can be one if there is only one group). After specifying the initial slice thickness and the slice range, the user is presented with an interactive display in which the control objects appear on the left and the slice visualization appears on the right. The user can change the orientation of the slice via the locator pane, which changes the projection matrix;  a slider controls the slice thickness; and there is an input field, which changes the slice center.
\textcolor{red}{insert sentence about new slice position display here.}
The user can also change the appearance of the plot by zooming into the center or changing the point size with the sliders provided. It is worth noting that this zoom works best with data that has been centered and scaled. The projected data can be displayed to contrast it with the slice by ticking a box. The explicit projection matrix can also be displayed via another checkbox. This is especially useful when we wish to import a projection that was identified as interesting in the manual exploration into a later study, as shown in the applications below. More details about the implementation and usage instructions are given in the Appendix.

@@

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardized) penguins data. The three species are reasonably different in size, with Gentoo distinguished from the other two on bill depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat + theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

@@

We start with a comparison of the two models and the data distribution shifting to $\mathbf{c}^{+}$, thus the slice is localized towards high values of bill depth in the top row of Figure \ref{slice1p}. \textcolor{blue}{The star glyph at the right indicates the slice center value, which shows that it is higher on `bd` and at the middle value for the other variables. This slice center guide is as proposed in @condviz2, and more details can be found in the Appendix.} We can see that all three slices (the two models and the data) contain almost no points from the third class (Gentoo) and that the decision boundary between the two models is very similar.

@@

Mathematica provided a useful sandbox to experiment with the ideas presented in this paper. 
Most of the data, though, is first constructed using R, especially the classification boundaries. \textcolor{red}{what should this say?}
If new technology for interactive graphics with precise cursor control becomes available in R it would be useful to create a tighter coupling of models and visualization to allow exploring and comparing fits.

@@

- The Mathematica source code defining the new functions in mmtour.wl.
- Three Mathematica notebooks with the code used in the application (corresponding to the three subsections).
- Appendix with additional details about manual controls, and Mathematica functions.
- Data files for the examples.
- R code to reproduce the example data, Figures 1 and 2 in the paper and Figures 1 and 2 in the appendix.
- Animations illustrating the manual tour and slicing, matching the static figures in the paper. These are also available at https://vimeo.com/747585410 and https://vimeo.com/747590472.
